---
title: "Biostat 212a Homework 1"
subtitle: "Due Jan 28, 2025 @ 11:59PM"
author: "Molly Shi (UID: 906558988)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    math: mathjax
    theme: cosmo
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
engine: knitr
knitr:
  opts_chunk: 
    fig.align: 'center'
    # fig.width: 6
    # fig.height: 4
    message: FALSE
    cache: false
editor: 
  markdown: 
    wrap: 72
---

## Filling gaps in lecture notes (10% pts)

Consider the regression model $$
Y = f(X) + \epsilon,
$$ where $\operatorname{E}(\epsilon) = 0$.

### Optimal regression function

Show that the choice $$
f_{\text{opt}}(X) = \operatorname{E}(Y | X)
$$ minimizes the mean squared prediction error $$
\operatorname{E}\{[Y - f(X)]^2\},
$$ where the expectations averages over variations in both $X$ and $Y$.
(Hint: condition on $X$.)

**Solution** The MSPE can be expressed as: $$ 
\begin{gather}
\mathbb{E}[(Y - f(X))^2] \\
= \mathbb{E}[(Y - f_{\text{opt}}(X) + f_{\text{opt}}(X) - f(X))^2] \\
= (Y - f_{\text{opt}}(X))^2 + (f_{\text{opt}}(X) - f(X))^2 + 2(Y - f_{\text{opt}}(X))(f_{\text{opt}}(X) - f(X))
\end{gather}
$$ Then, we take the expectation of both sides: $$
\begin{gather}
\mathbb{E}[(Y - f(X))^2] \\
= \mathbb{E}[(Y - f_{\text{opt}}(X))^2] + \mathbb{E}[(f_{\text{opt}}(X) - f(X))^2] + 2\mathbb{E}[(Y - f_{\text{opt}}(X))(f_{\text{opt}}(X) - f(X))]
\end{gather}
$$Conditioning on $X$ and using the fact that
$f_{\text{opt}}(X) = \mathbb{E}[Y|X]$, the term $Y - f_{\text{opt}}(X)$
is equal to 0, which turns the equation into:

$$
\mathbb{E}[(Y - f(X))^2] = \mathbb{E}[(Y - f_{\text{opt}}(X))^2] + \mathbb{E}[(f_{\text{opt}}(X) - f(X))^2]
$$

Since $\mathbb{E}[(f_{\text{opt}}(X) - f(X))^2] \geq 0$, the MSPE is
minimized when $f(X) = f_{\text{opt}}(X)$, which is
$$f(X) = \mathbb{E}[Y|X].$$

### Bias-variance trade-off

Given an estimate $\hat f$ of $f$, show that the test error at a $x_0$
can be decomposed as $$
\operatorname{E}\{[y_0 - \hat f(x_0)]^2\} = \underbrace{\operatorname{Var}(\hat f(x_0)) + [\operatorname{Bias}(\hat f(x_0))]^2}_{\text{MSE of } \hat f(x_0) \text{ for estimating } f(x_0)} + \underbrace{\operatorname{Var}(\epsilon)}_{\text{irreducible}},
$$ where the expectation averages over the variability in $y_0$ and
$\hat f$.

**Solution** To decompose$\operatorname{E}\{[y_0 - \hat f(x_0)]^2\}$
into
$\text{Var}(\hat{f}(x_0)) + [\text{Bias}(\hat{f}(x_0))]^2 + \text{Var}(\epsilon)$,
we can first express the test error by substituting
$y_0 = f(x_0) + \epsilon$ :

$$
\begin{gather}
\mathbb{E}[(y_0 - \hat{f}(x_0))^2] \\
= \mathbb{E}[(f(x_0) + \epsilon - \hat{f}(x_0))^2] \\
= \mathbb{E}[(f(x_0) - \hat{f}(x_0))^2] + 2\mathbb{E}[(f(x_0) - \hat{f}(x_0))\epsilon] + \mathbb{E}[\epsilon^2]
\end{gather}
$$

We can then simplify the terms based on $\epsilon$ is independent of
$\hat{f}(x_0)$ and has zero mean, which makes
$\mathbb{E}[(f(x_0) - \hat{f}(x_0))\epsilon] = 0$

Thus, we can simplify the equation as

$$
\begin{gather}
\mathbb{E}[(y_0 - \hat{f}(x_0))^2] \\
= \mathbb{E}[(f(x_0) - \hat{f}(x_0))^2] + \mathbb{E}[\epsilon^2]
\end{gather}
$$

Then, to further decompose $\mathbb{E}[(f(x_0) - \hat{f}(x_0))^2]$, we
subtract and add term $\mathbb{E}[\hat{f}(x_0)]$ and expand the equation
as:

$$
\begin{gather}
(f(x_0) - \hat{f}(x_0))^2 \\
= (f(x_0) - \mathbb{E}[\hat{f}(x_0)])^2 + (\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0))^2 + 2(f(x_0) - \mathbb{E}[\hat{f}(x_0)])(\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0))
\end{gather}
$$

In this case, the third term
$\mathbb{E}[\hat{f}(x_0)])(\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0)) = 0$
because $\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0)$ as the deviation of
the estimator $\hat{f}(x_0)$ from its expectation, will have a zero mean
by definition of variance.

Therefore, the equation will be:

$$
\begin{gather}
\mathbb{E}[(y_0 - \hat{f}(x_0))^2] \\
= \mathbb{E}[(f(x_0) - \hat{f}(x_0))^2] + \mathbb{E}[\epsilon^2] \\
= \mathbb{E}(f(x_0) - \mathbb{E}[\hat{f}(x_0)])^2 + \mathbb{E}(\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0))^2 + \mathbb{E}[\epsilon^2] 
\end{gather}
$$

where first term becomes the bias squared
$\mathbb{E}[(f(x_0) - \mathbb{E}[\hat{f}(x_0)])^2] = [\text{Bias}(\hat{f}(x_0))]^2$,

and the second term becomes the variance
$\mathbb{E}[(\mathbb{E}[\hat{f}(x_0)] - \hat{f}(x_0))^2] = \text{Var}(\hat{f}(x_0))$,

and the third term $\mathbb{E}[\epsilon^2]$ is the irreducible error.

Therefore, we can conclude that $$
\operatorname{E}\{[y_0 - \hat f(x_0)]^2\} = \text{Var}(\hat{f}(x_0)) + [\text{Bias}(\hat{f}(x_0))]^2 + \text{Var}(\epsilon)
$$

## ISL Exercise 2.4.3 (10% pts)

We now revisit the bias-variance decomposition.

\(a\) Provide a sketch of typical (squared) bias, variance, training
error, test error, and Bayes (or irreducible) error curves, on a single
plot, as we go from less flexible statistical learning methods towards
more flexible approaches. The x-axis should represent the amount of
flexibility in the method, and the y-axis should represent the values
for each curve. There should be five curves. Make sure to label each
one.

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Define the x values
x <- seq(0, 10, by = 0.02)

# Define the functions
squared_bias <- function(x) {
  0.002 * (-x + 10)^3
}
variance <- function(x) {
  0.002 * x^3
}
training_error <- function(x) {
  2.38936 - 0.825077 * x + 0.176655 * x^2 - 0.0182319 * x^3 + 0.00067091 * x^4
}
test_error <- function(x) {
  3 - 0.6 * x + 0.06 * x^2
}
bayes_error <- function(x) {
  x + 1 - x
}

# Create a data frame with all values
data <- tibble(
  x = x,
  `squared bias` = squared_bias(x),
  variance = variance(x),
  `training error` = training_error(x),
  `test error` = test_error(x),
  `Bayes error` = bayes_error(x)
) %>%
  pivot_longer(cols = -x, names_to = "Metric", values_to = "Value")

# Plot using ggplot2
ggplot(data, aes(x = x, y = Value, color = Metric)) +
  geom_line(size = 1) +
  labs(
    x = "Model Flexibility",
    y = "Value",
    title = "Bias-Variance Tradeoff Plot",
    color = "Metric"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top"
  )
```

```{r, eval = F}
library(tidyverse)
# Advertising
Advertising <- read_csv("../hw1/Advertising.csv",  col_select = TV:sales) %>% 
  print(width = Inf)

fit <- lm(sales ~ TV, data = Advertising)

```

\(b\) Explain why each of the five curves has the shape displayed in
part (a).

**Solution**

1.  Squared bias will decreases with increasing flexibility because more
    flexible methods will have less bias.

2.  Variance will increase with increasing flexibility because more
    flexible methods will have higher variance.

3.  Training error will decrease with increasing flexibility because
    more complex model will have better fit over training data.

4.  Test error will decrease at first but increase overtime because more
    flexibility will induce overfitting, which means the model has a
    better fit over training data over testing data.

5.  Irreducible error does not change with model flexibility.

## ISL Exercise 2.4.4 (10% pts)

You will now think of some real-life applications for statistical
learning.

\(a\) Describe three real-life applications in which classification
might be useful. Describe the response, as well as the predictors. Is
the goal of each application inference or prediction? Explain your
answer.

**Solution**

| Real-life application | Predictors | Goal (inference/prediction) |
|----|----|----|
| Image classification | Pixel intensities of the image | Prediction |
| Virus type classification | Genetic sequences, patient symptoms, geographic data | Prediction |
| Spam email detection | Word frequencies, sender address, keywords, metadata | Prediction |

\(b\) Describe three real-life applications in which regression might be
useful. Describe the response, as well as the predictors. Is the goal of
each application inference or prediction? Explain your answer.

**Solution**

| Real-life application | Predictors | Goal (inference/prediction) |
|----|----|----|
| Patient's health monitoring | Age, diet, physical activity, medication | Prediction |
| Predicting student's score | Study hours, attendance, prior grades, resources used | Prediction |
| Profit prediction | Marketing budget, product price, customer demographics, economic indicators | Inference |

\(c\) Describe three real-life applications in which cluster analysis
might be useful.

**Solution**

1)  Gene Expression Clustering: Grouping genes with similar expression
    profiles to identify functional relationships or biological
    pathways.
2)  Drug Discovery: Identifying clusters of chemical compounds with
    similar molecular structures or pharmacological properties for new
    drug development.
3)  Customer Segmentation: Grouping customers based on purchasing
    behaviors, demographics, or preferences to tailor marketing
    strategies.

## ISL Exercise 2.4.10 (30% pts)

This exercise involves the `Boston` housing data set.

(a) To begin, load in the Boston data set. The Boston data set is part
    of the ISLR2 library in R. How many rows are in this data set? How
    many columns? What do the rows and columns represent?

Your can read in the `boston` data set directly from url
<https://raw.githubusercontent.com/ucla-biostat-212a/2024winter/master/slides/data/Boston.csv>.
A documentation of the `boston` data set is
[here](https://www.rdocumentation.org/packages/ISLR2/versions/1.3-2/topics/Boston).

::: panel-tabset
#### R

```{r, evalue = F}
library(tidyverse)
library(ISLR2)

Boston <- read_csv("https://raw.githubusercontent.com/ucla-biostat-212a/2024winter/master/slides/data/Boston.csv", col_select = -1) %>% 
  print(width = Inf)

Boston
```

#### Python

```{python}
import pandas as pd
import io
import requests

url = "https://raw.githubusercontent.com/ucla-econ-425t/2023winter/master/slides/data/Boston.csv"
s = requests.get(url).content
Boston = pd.read_csv(io.StringIO(s.decode('utf-8')), index_col = 0)
Boston
```
:::

**Solution** There are 506 rows and 13 columns in this dataset. Each row
in this dataset represent a house, and each column represent a property
of the house. Description of the data columns are:

-   `CRIM` - per capita crime rate by town

-   `ZN` - proportion of residential land zoned for lots over 25,000
    sq.ft.

-   `INDUS` - proportion of non-retail business acres per town.

-   `CHAS` - Charles River dummy variable (1 if tract bounds river; 0
    otherwise)

-   `NOX` - nitric oxides concentration (parts per 10 million)

-   `RM` - average number of rooms per dwelling

-   `AGE` - proportion of owner-occupied units built prior to 1940

-   `DIS` - weighted distances to five Boston employment centres

-   `RAD` - index of accessibility to radial highways

-   `TAX` - full-value property-tax rate per \$10,000

-   `PTRATIO` - pupil-teacher ratio by town

-   `LSTAT` - % lower status of the population

-   `MEDV` - Median value of owner-occupied homes in \$1000's

b.  Make some pairwise scatterplots of the predictors (columns) in this
    data set. Describe your findings.

    ```{r}
    library(ggplot2)

    ggplot(Boston, aes(tax, rm)) + 
      geom_point()
    ggplot(Boston, aes(indus, crim)) +
      geom_point()
    ```

    **Solution** From the scatterplot between `tax` (full-value
    property-tax rate per \$10,000) and `rm` (average number of rooms),
    we can observe that the house that have less property-tax rate per
    \$10,000 usually have more rooms. And from the scatterplot between
    `crim` (crime rate) and `indus` (proportion of non-retail business
    per acres per town), we can see that the high crime rate are
    associated with a particular `indus` level (around 17-18), which is
    quite interesting.

c.  Are any of the predictors associated with per capita crime rate? If
    so, explain the relationship.

    ```{r}
    heatmap(cor(Boston, method = "spearman"), cexRow = 1.1, cexCol = 1.1)
    ```

    **Solution** Yes. From the heat map with Spearman correlation, we
    can observe that `crim` have the highest association with `chas`
    (the Charles River dummy variable), which indicate some correlation
    between crime rate and adjacency to Charles River. In addition,
    `crim` also have relative high correlation coefficient with `dis`,
    `rm`, `zn`, `ptratio` .

d.  Do any of the census tracts of Boston appear to have particularly
    high crime rates? Tax rates? Pupil-teacher ratios? Comment on the
    range of each predictor.

    ```{r}
    Boston |>
      pivot_longer(cols = 1:13) |>
      filter(name %in% c("crim", "tax", "ptratio")) |>
      ggplot(aes(value)) +
      geom_histogram(bins = 20) +
      facet_wrap(~name, scales = "free", ncol = 1)
    ```

    **Solution**

    -   The **`crim`** variable is highly skewed to the right, with many
        outliers at the upper end of the distribution. This suggests
        that the majority of towns have very low crime rates, likely
        between 0 and 5, while a few towns exhibit extremely high crime
        rates exceeding 70. Overall, the crime rate ranges from 0 to
        approximately 80.
    -   The **`ptratio`** variable shows a relatively mild right skew.
        The values range from 12.5 to 22, with no significant outliers
        observed. This suggests a more uniform distribution of
        pupil-teacher ratios among the towns.
    -   The **`tax`** variable is skewed to the left, indicating that
        most towns have lower property-tax rates. The distribution
        ranges from 200 to 800, but there is a noticeable gap between
        500 and 600, where no towns have tax rates in that interval.

e.  How many of the census tracts in this data set bound the Charles
    river?

    ```{r}
    table(Boston$chas)
    ```

f.  What is the median pupil-teacher ratio among the towns in this data
    set?

    ```{r}
    median(Boston$ptratio)
    ```

g.  Which census tract of Boston has lowest median value of
    owner-occupied homes? What are the values of the other predictors
    for that census tract, and how do those values compare to the
    overall ranges for those predictors? Comment on your findings.

    ```{r}
    Boston <- data.frame("obs" = c(1:length(Boston$crim)), Boston)
    Boston %>%  filter(medv == min(medv))
    ```

    **Solution** The The above two census tracts have the lowest median
    value of 5 from the owner occupied homes.

    ```{r}
    summary(Boston[,2:14])
    ```

    -   `crim`: The crime rates in both tracts are outliers,
        significantly higher than the median crime rate.
    -   `zn`: Both tracts have zero proportion of land zoned for lots
        over 25,000 sq. ft., which matches the median value, suggesting
        minimal or no investment in these areas.
    -   `indus`: The proportion of non-retail business acres is 18.1 for
        both tracts, exceeding the median and falling in the third
        quartile, indicating promising business opportunities.
    -   `chas`: Neither tract is adjacent to the Charles River.
    -   `nox`: Nitrogen oxide concentrations are above the third
        quartile, suggesting high levels of pollution, potentially due
        to proximity to highways.
    -   `rm`: The average number of rooms per dwelling falls in the
        first quartile, indicating smaller homes.
    -   `age`: The proportion of owner-occupied units built before 1940
        is in the upper quartile, reflecting very old housing stock.
    -   `dis`: The weighted average distance to five Boston employment
        centers is in the lower quartile, possibly indicating high
        unemployment.
    -   `rad`: Both tracts have the maximum index for radial highway
        accessibility, implying close proximity to highways.
    -   `tax`: Property tax rates per \$10,000 are near the third
        quartile, indicating high taxes for smaller homes, suggesting a
        non-linear relationship between tax rates and housing size.
    -   `ptratio`: The pupil-teacher ratio is near the third quartile,
        suggesting a relatively strong education system.
    -   `lstat`: The proportion of the population in lower socioeconomic
        status is in the upper quartile, indicating a high concentration
        of low-income households.

h.  In this data set, how many of the census tract average more than
    seven rooms per dwelling? More than eight rooms per dwelling?
    Comment on the census tracts that average more than eight rooms per
    dwelling.

    ```{r}
    seven_rooms_above = Boston %>% filter(rm > 7)
    eight_rooms_above = Boston %>% filter(rm > 8)
    data.frame("more_than_7_rooms" = c(length(seven_rooms_above$rm)),
               "more_than_8_rooms" = c(length(eight_rooms_above$rm)))
    ```

    ```{r}
    plot_all = boxplot(Boston[, 2:14], main="All House Distribution")
    plot_eight = boxplot(eight_rooms_above[, 2:14], main="House > 8 Bedroom Distribution")
    ```

    **Solution** Comparing the box plot distribution between all houses
    and houses with more than eight bedrooms, the tax is generally lower
    and the crime rate are also lower for houses with more than 8
    bedrooms.

## ISL Exercise 3.7.3 (20% pts)

Suppose we have a data set with five predictors, $X_1$ = GPA, $X_2$ =
IQ, $X_3$ = Level (1 for College and 0 for High School), $X_4$ =
Interaction between GPA and IQ, and $X_5$ = Interaction between GPA and
Level. The response is starting salary after graduation (in thousands of
dollars). Suppose we use least squares to fit the model, and get
$\hat\beta_0 = 50$, $\hat\beta_1 = 20$, $\hat\beta_2 = 0.07$,
$\hat\beta_3 = 35$, $\hat\beta_4 =
0.01$, $\hat\beta_5 = -10$.

a.  Which answer is correct, and why?

    i.  For a fixed value of IQ and GPA, high school graduates earn more
        on average than college graduates.

    ii. For a fixed value of IQ and GPA, college graduates earn more on
        average than high school graduates.

    iii. For a fixed value of IQ and GPA, high school graduates earn
         more on average than college graduates provided that the GPA is
         high enough.

    iv. For a fixed value of IQ and GPA, college graduates earn more on
        average than high school graduates provided that the GPA is high
        enough.

    **Solution** According to the given information, the model is

    $$y = \beta_0 + \beta_1 \cdot \text{GPA} + \beta_2 \cdot \text{IQ} + \beta_3 \cdot \text{Level} + \beta_4 \cdot \text{GPA} \cdot \text{IQ} + \beta_5 \cdot \text{GPA} \cdot \text{Level}$$

    For a fixed IQ and GPA, we can calculate the difference in salary by

    $$
    \Delta y = y_{level=1} - y_{level=0} = \beta_3 + \beta_5 \cdot \text{GPA}
    $$

    Since $\Delta y > 0$ , we can derive that

    $$
    \beta_3 + \beta_5 \cdot \text{GPA} > 0 \\
    \text{GPA} < \dfrac{-\beta_3}{\beta_5} = \dfrac{-35}{-10} = 3.5
    $$

    Hence, **option iii** is correct: for a fixed value of IQ and GPA,
    high school graduates earn more, on average, than college graduates
    provided that the GPA is more than equal to 3.5.

b.  Predict the salary of a college graduate with IQ of 110 and a GPA \>
    of 4.0.

    ```{r}
    model <- function(gpa, iq, level) {
      50 +
        gpa * 20 +
        iq * 0.07 +
        level * 35 +
        gpa * iq * 0.01 +
        gpa * level * -10
    }
    x <- seq(1, 5, length = 10)
    y <- seq(1, 200, length = 20)
    college <- t(outer(x, y, model, level = 1))
    high_school <- t(outer(x, y, model, level = 0))

    model(gpa = 4, iq = 110, level = 1)
    ```

    **Solution** The salary of a college graduate with IQ of 110 and a
    GPA \> of 4.0 is \$137.1 k.

c.  True or false: Since the coefficient for the GPA/IQ interaction term
    is very small, there is very little evidence of an interaction
    effect. Justify your answer.

    **Solution** False. It's because GPA and IQ operate on different
    scales. A more effective approach would be to explicitly test the
    significance of the interaction effect and/or visualize or quantify
    its impact on sales within realistic ranges of GPA and IQ values.

## ISL Exercise 3.7.15 (20% pts)

This problem involves the `Boston` data set, which we saw in the lab for
this chapter. We will now try to predict per capita crime rate using the
other variables in this data set. In other words, per capita crime rate
is the response, and the other variables are the predictors.

a.  For each predictor, fit a simple linear regression model to predict
    the response. Describe your results. In which of the models is there
    a statistically significant association between the predictor and
    the response? Create some plots to back up your assertions.

    ```{r}
    pred <- subset(Boston, select = -crim)
    fits <- lapply(pred, function(x) lm(Boston$crim ~ x))
    printCoefmat(do.call(rbind, lapply(fits, function(x) coef(summary(x))[2, ])))
    ```

    **Solution** All the estimate values can be used as predictors for
    variable "crim," except for the independent variable `chas`, as
    there appears to be no association between tracts near the Charles
    River and crime rate in the area. Most variables are individually
    significant as predictors.

b.  Fit a multiple regression model to predict the response using all of
    the predictors. Describe your results. For which predictors can we
    reject the null hypothesis $H_0 : \beta_j = 0$?

    ```{r}
    mfit <- lm(crim ~ ., data = Boston)
    summary(mfit)
    ```

    **Solution** At a significance level of 0.05, variable `zn`, `dis`,
    `rad`, `black` and `medv` have significant association with `crim`.
    Collectively, the predictors account for 44.93% of the variation in
    the crime rate. This suggests that using all variables together
    explains a greater proportion of the variation compared to simple
    regression, where the highest variance explained was 39%. The full
    model improves the explained variation by an additional 5.9%.

c.  How do your results from (a) compare to your results from (b)?
    Create a plot displaying the univariate regression coefficients
    from (a) on the $x$-axis, and the multiple regression coefficients
    from (b) on the $y$-axis. That is, each predictor is displayed as a
    single point in the plot. Its coefficient in a simple linear
    regression model is shown on the x-axis, and its coefficient
    estimate in the multiple linear regression model is shown on the
    y-axis.

    ```{r}
    plot(sapply(fits, function(x) coef(x)[2]), coef(mfit)[-1],
      xlab = "Univariate regression",
      ylab = "multiple regression"
    )
    ```

    **Solution** Based on the plot, we can observe that the fitting
    from (b) show reduced significance compared to the models in part
    (a). The most notable difference in coefficient values between the
    simple and multiple models is observed for `nox` In the simple
    model, the coefficient for `nox` is 31.25, while in the multiple
    model, it is -9.96. Coefficients for other predictors also show
    differences between the two models.

d.  Is there evidence of non-linear association between any of the
    predictors and the response? To answer this question, for each
    predictor X, fit a model of the form $$
    Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \epsilon
    $$

    ```{r}
    pred <- subset(pred, select = -chas)
    fits <- lapply(names(pred), function(p) {
      f <- paste0("crim ~ poly(", p, ", 3)")
      lm(as.formula(f), data = Boston)
    })
    for (fit in fits) printCoefmat(coef(summary(fit)))
    ```

    **Solution** Based on the above summary plot, there are many
    variables which have non-linearity. For variable `zn`, `rm`, `rad`,
    `tax`, and `lstat`, the squared coefficient is significant. For
    variable `indus`, `nox`, `age`, `dis`, `ptratio` and `medv` , the
    quadratic coefficient is significant.

## Bonus question (20% pts)

For multiple linear regression, show that $R^2$ is equal to the
correlation between the response vector
$\mathbf{y} = (y_1, \ldots, y_n)^T$ and the fitted values
$\hat{\mathbf{y}} = (\hat y_1, \ldots, \hat y_n)^T$. That is $$
R^2 = 1 - \frac{\text{RSS}}{\text{TSS}} = [\operatorname{Cor}(\mathbf{y}, \hat{\mathbf{y}})]^2
$$**Solution** In multiple linear regression, the **RSS** (residual sum
of squares) and **TSS**(total sum of squares), and **ESS** (explained
sum of square) can be written as:

$$
\begin{gather}
\text{RSS} = \sum_{i=1}^n (y_i - \hat{y}_i)^2 \\
\text{TSS} = \sum_{i=1}^n (y_i - \bar{y})^2 \\
\text{ESS} = \sum_{i=1}^n (\hat{y}_i - \bar{y})^2 \\
and\:\: \text{TSS} = \text{ESS}+\text{RSS}
\end{gather}
$$

Note that **TSS** can be decomposed into the sum of **ESS** and **RSS**
because the observed $y$ is split into the fitted values $\hat{y}$​ and
the residuals.

Alternatively, $R^2$ measures the proportion of the total variability in
$\mathbf{y}$ that is explained by the regression model, which can also
be written as:

$$
\begin{gather}
R^2 = 1 - \frac{\text{RSS}}{\text{TSS}} \\
= \frac{\text{TSS}}{\text{TSS}} - \frac{\text{RSS}}{\text{TSS}} \\ 
= \frac{\text{ESS}}{\text{TSS}}
\end{gather}
$$

On the other hand, the squared correlation can be decomposed as
covariance over variance:

$$
[\operatorname{Cor}(\mathbf{y}, \hat{\mathbf{y}})]^2 = \frac{\operatorname{Cov}(\mathbf{y}, \hat{\mathbf{y}})^2}{\operatorname{Var}(\mathbf{y}) \cdot \operatorname{Var}(\hat{\mathbf{y}})}
$$

The covariance between $\mathbf{y}$ and $\hat{\mathbf{y}}$ , in this
case, $\operatorname{Cov}(\mathbf{y}, \hat{\mathbf{y}})$ can be written
as $\operatorname{Var}(\hat{\mathbf{y}})$ since the fitted values
$\hat{\mathbf{y}}$ are linear combinations of $\mathbf{y}$
($\hat{\mathbf{y}}=X\hat{\mathbf{\beta}}$

$$
\begin{gather}
\operatorname{Cov}(\mathbf{y}, \hat{\mathbf{y}}) = \frac{1}{n} \sum_{i=1}^n (y_i - \bar{y})(\hat{y}_i - \bar{y}) \\
 = \operatorname{Var}(\hat{\mathbf{y}}) = \frac{\text{ESS}}{n}
\end{gather}
$$

Given that:

$$
\operatorname{Cov}(\mathbf{y}, \hat{\mathbf{y}}) = \frac{\text{ESS}}{n}, \operatorname{Var}(\mathbf{y}) = \frac{\text{TSS}}{n}, \operatorname{Var}(\hat{\mathbf{y}}) = \frac{\text{ESS}}{n}
$$

We can substitute covariance and variance in terms of:

$$
\begin{gather}
[\operatorname{Cor}(\mathbf{y}, \hat{\mathbf{y}})]^2 = \frac{\operatorname{Cov}(\mathbf{y}, \hat{\mathbf{y}})^2}{\operatorname{Var}(\mathbf{y}) \cdot \operatorname{Var}(\hat{\mathbf{y}})}\\
= \frac{\left(\frac{\text{ESS}}{n}\right)^2}{\left(\frac{\text{TSS}}{n}\right) \cdot \left(\frac{\text{ESS}}{n}\right)} \\
= \frac{\text{ESS}}{\text{TSS}}
\end{gather}
$$

Since $R^2 = \frac{\text{ESS}}{\text{TSS}}$ , we can conclude that
$R^2 = 1 - \frac{\text{RSS}}{\text{TSS}} = [\operatorname{Cor}(\mathbf{y}, \hat{\mathbf{y}})]^2$
